---
title: "p8131_stl2137_hw5"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(pscl)
```

## Do we do intercepts?

# Problem 1

```{r}
### Load in crab data

crab_dat <- read.table("/Users/SabrinaLin/Documents/Spring_20_Biostat_Methods_2/Homework/p8131_stl2137_hw5/HW5-crab.txt", header = TRUE) %>% 
  janitor::clean_names()
```

## Part A

```{r}
### M1 model
crab_m1_mod <- glm(sa ~ w, family = poisson(link = "log"), data = crab_dat)
summary(crab_m1_mod)

### Goodness of fit
crab_m1_pval <- pchisq(crab_m1_mod$deviance, df = crab_m1_mod$df.residual)
```

* 0.16405 is the log rate ratio of number of satellites with every 1 unit increase in carapace width. 

* With a p-value of `r crab_m1_pval`, we reject the null hypothesis and conclude that the model does not fit the data.  

## Part B

```{r}
### M2 model
crab_m2_mod <- glm(sa ~ w + wt, family = poisson(link = "log"), data = crab_dat)
summary(crab_m2_mod)

### M2 vs M1
test.stat = crab_m1_mod$deviance - crab_m2_mod$deviance
df = 171 - 170
pval = 1 - pchisq(test.stat, df = df)
pval
```

* With a p-value of 0.32640 compared to $\alpha = 0.05$, carapace width is an insignifcant variable. 0.04590 is the log rate ratio of number of satellites with every 1 unit increase in carapace width, keeping weight constant. 

* 0.44744 is the log rate ratio of the number of satellites with every 1 unit increase in weight, keeping carapace width constant. 

### Goodness of Fit: Which model is better?

* With a p-value of `r pval`, we reject the null hypothesis and conclude that M2, the bigger model, is the model we should continue with. 

## Part C

```{r}
m2_dispersion_mod <- glm(sa ~ w + wt, family = quasi(link = "log", variance = mu), data = crab_dat)
summary(m2_dispersion_mod)

## Checking that traditional way also works
res.m2.disp <- residuals(crab_m2_mod, type = 'pearson', data = crab_dat)
G1.m2 <- sum(res.m2.disp^2)
pval = 1 - pchisq(G1.m2, df = 170)
phi = G1.m2/170
#phi

#summary(crab_m2_mod , dispersion = phi)

```

* With a p-value of 0.581 compared to $\alpha = 0.05$, the carapace width variable is insignificant. 0.04590 is the log rate ratio of number of satellites with every 1 unit increase in carapace width, holding weight constant. 

* With a p-value of 0.114 compared to $\alpha = 0.05$, the weight variable is insignificant. 0.44744 is the log rato ratio of number of satellites with every 1 unit increase in weight, holding carapace width constant. 

```{r}
### Half normal plot
tilde.phi = crab_m2_mod$deviance/crab_m2_mod$df.residual
res = residuals(crab_m2_mod, type = 'pearson')
plot(qnorm((173+1:173+0.5)/(2*173+1.125)),sort(abs(res)),xlab='Expected Half-Normal Order Stats',ylab='Ordered Abs Pearson Residuals')
abline(a=0,b=1)
abline(a=0,b=sqrt(phi),lty=2)
```

As demonstrated by the plot, the data does not follow a $\phi = 1$, showing that there is overdispersion. 

# Problem 2

```{r}
### Load parasite data

parasite_dat <- read.table("/Users/SabrinaLin/Documents/Spring_20_Biostat_Methods_2/Homework/p8131_stl2137_hw5/HW5-parasite.txt", header = TRUE) %>% 
  janitor::clean_names() %>% 
  mutate(
    area = as.factor(area),
    year = as.factor(year)
  )
```

## Part A

```{r}
parasite_mod <- glm(intensity ~ area + year + length, family = poisson(link = "log"), data = parasite_dat)

summary(parasite_mod)
```

* `r parasite_mod$coefficients[2]` is the log rate ratio of the number of parasites when going from area 2 to area 1, holding all other variables constant. 

* `r parasite_mod$coefficients[3]` is the log rate ratio of the number of parasites when going from area 3 to area 1, holding all other variables constant. 

* `r parasite_mod$coefficients[4]` is the log rate ratio of the number of parasites when going from area 4 to area 1, holding all other variables constant. 

* `r parasite_mod$coefficients[5]` is the log rate ratio of the number of parasites when going from the year 2000 to the year 1999, holding all other variables constant.

* `r parasite_mod$coefficients[6]` is the log rate ratio of the number of parasites when going from the year 2001 to the year 1999, holding all other variables constant.

* `r parasite_mod$coefficients[7]` is the log rate ratio of the number of parasites for every 1 unit increase in length, holding all other variables constant.

## Part B

```{r}
parasite_pval <- pchisq(parasite_mod$deviance, df = parasite_mod$df.residual)
```
With a p-value of `r parasite_pval`, we reject the null hypothesis and conclude that the model does not fit the data. 

## Part C

```{r}
### what should the given be? area? length? 

parasite_zero_mod <- zeroinfl(intensity ~ year + length | area, data = parasite_dat)
summary(parasite_zero_mod)
```

### Count Model Coefficient Interpretations

* `r parasite_zero_mod$coefficients$count[2]` is the log rate ratio of the number of parasites when going from the year 2000 to the year 1999 given that the fish strain is susceptible to parasites, holding all other variables constant. 

* `r parasite_zero_mod$coefficients$count[3]` is the log rate ratio of the number of parasites when going from the year 2001 to the year 1999 given that the fish strain is susceptible to parasites, holding all other variables constant. 

* `r parasite_zero_mod$coefficients$count[4]` is the log rate ratio of the number of parasites for every 1 unit increase in length given that the fish strain is susceptible to parasites, holding all other variables constant.

### Zero-inflation Model Coefficient Interpretations

* `r parasite_zero_mod$coefficients$zero[2]` is the log odds ratio of not being the fish strain susceptible to parasites when comparing area 2 to area 1. 

* `r parasite_zero_mod$coefficients$zero[3]` is the log odds ratio of not being the fish strain susceptible to parasites when comparing area 3 to area 1. 

* `r parasite_zero_mod$coefficients$zero[4]` is the log odds ratio of not being the fish strain susceptible to parasites when comparing area 4 to area 1. 

